
\documentclass{mcom-l}


\begin{document}
\title{On the development of a computer algebra module for Quandle Theory}

%    Information for first author
\author{Niccolò Carrivale - s1035576}

\maketitle


Computer algebra involves developing algorithms to carry out exact computations involving algebraic structures. Quandles are algebraic structures that can benefit from the help of computational tools and attention from the community, much like groups in the past. This project introduces a modern tool able to work with quandles: congruences, subuniverses, isomorphisms, and monomorphisms, among others; built on the solid mathematical foundations of the most recent developments in Quandle Theory, and Universal Algebra as well as computational tools that withstood the test of time such as Magma and Rig, it is able to obtain useful results at a competitive speed. As an example, it will be shown how to compute and visualise connections among connected quandles. 

L'algebra computazionale si occupa di sviluppare algoritmi esatti per facilitare le elaborazioni simboliche con strutture algebriche. I quandles sono strutture algebriche che possono beneficiare dall'aiuto di strumenti computazionali e dall'attenzione della comunità, come i gruppi in passato. Questo progetto presenta uno strumento moderno in grado di lavorare con i quandles: congruenze, subuniversi, isomorfismi e monomorfismi, tra le altre cose; siluppato sulle solide fondamenta matematiche dei più recenti sviluppi nella teoria dei quandles e algebra universale così come strumenti computazionali che hanno resistito alla prova del tempo come Magma e Rig, è in grado di ottenere risultati utili a velocità competitive. Come esempio, verrà mostrato come calcolare e visualizzare le connessioni tra quandle connessi.



\section{Introduction}
\subfile{chapters/Introduction}

\section{Introduction to Quandles}
\subfile{chapters/IntroductionQuandles}

\section{Quandle Homomorphisms}
\subfile{chapters/QuandleHomomorphisms}


\bibliographystyle{amsalpha}
\bibliography{references.bib}



\end{document}


